Process the data by review to then use for model selection and prediction
========================

Load and explore the data
```{r setup}
## Libraries
library(plyr)
library(doMC)

## Set up for using plyr in parallel
registerDoMC(20)
getDoParWorkers()

## Read data
load("../processReview/testProcessed.Rdata")
load("../processReview/trainingProcessed.Rdata")

## Explore it
#lapply(testProcessed, head)
lapply(testProcessed, summary)

#lapply(trainingProcessed, head)
lapply(trainingProcessed, summary)
```

Merge the data just to be able to then process the data properly. I mean, some businesses' might not be defined in the test data and you have to find them in the training data. At least, I think that's what some were discussing in the forums.

```{r merge}
types <- names(trainingProcessed)
names(types) <- types
all <- lapply(types, function(x) {
	## For lazy code
	y <- trainingProcessed[[x]]
	z <- testProcessed[[x]]
	
	## Find if the columns match
	toadd <- colnames(y)[!colnames(y) %in% colnames(z)]
	new <- z
	
	## Add empty columns if necessary (aka, for the votes columns not present in the test data)
	if(length(toadd) > 0) {
		empty <- data.frame(matrix(NA, ncol=length(toadd), nrow=nrow(z)))
		colnames(empty) <- toadd
		new <- cbind(z, empty)
	}
	
	## Reorder the columns so the rbind makes sense
	idx <- sapply(colnames(y), function(m) { which(colnames(new) == m) })
	new <- new[, idx]
	
	## Finish
	rbind(y, new)
})

## Check that it makes sense
identical(unlist(lapply(all, nrow)), unlist(lapply(types, function(x) { nrow(testProcessed[[x]]) + nrow(trainingProcessed[[x]])})))
```

Define the function that will process the information for the reviews. The goal is to pull the information from the other tables (business, checkin, user) into a big review table. Then this table can be the input for the modelling functions in R.

```{r processFun}
## Helper functions for ldply
processUser <- function(x, user.a) {
	i <- which(user.a$user_id == x)
	which.cols <- colnames(user.a) %in% c("name", "average_stars", "review_count")
	if(length(i) >= 1) {
		res <- user.a[i, which.cols]
	} else {
		res <- user.a[i, which.cols]
		res[1, ] <- rep(NA, length(which.cols))
	} 
	if(length(i) > 1) {
		warning("Unexpected case. Only using the info from the first match.")
		print(user.a[i, which.cols])
	}
	## Avoid variable names overlap with columns from other tables
	colnames(res) <- paste0("user.", colnames(res))
	return(res)
}

processCheckin <- function(x, checkin.a, small) {
	i <- which(checkin.a$business_id == x)
	if(small){
		which.cols <- colnames(checkin.a) %in% c("checkin_info.total", "type")
	} else {
		which.cols <- !colnames(checkin.a) %in% c("business_id")
	}	
	if(length(i) >= 1){
		res <- checkin.a[i, which.cols]
	} else {
		res <- checkin.a[i, which.cols]
		res[1, ] <- rep(NA, length(which.cols))
	}
	if(length(i) > 1){
		warning("Unexpected case. Only using the info from the first match.")
		print(checkin.a[i, which.cols])
	}
	## Avoid variable names overlap with columns from other tables
	colnames(res) <- gsub("_info", "", colnames(res))
	## Workaround to avoid issues with vectors and data.frames
	res <- res[, -which(colnames(res)=="type")]
	return(res)
}

processBusiness <- function(x, business.a) {
	i <- which(business.a$business_id == x)
	## Get the zipcode
	business.a$zipcode <- as.integer(gsub("^.*(AZ|CA|CO|SC)(| )", "", business.a$full_address))
	
	which.cols <- !colnames(business.a) %in% c("business_id", "state", "type", "full_address", "neighborhoods", "name")
	if(length(i) >= 1){
		res <- business.a[i, which.cols]
	} else {
		res <- business.a[i, which.cols]
		res[1, ] <- rep(NA, length(which.cols))
	}
	if(length(i) > 1){
		warning("Unexpected case. Only using the info from the first match.")
		print(business.a[i, which.cols])
	}
	## Avoid variable names overlap with columns from other tables
	colnames(res) <- paste0("business.", colnames(res))
	return(res)
}

processFun <- function(input, all, parallel=TRUE, small=FALSE) {
	
	## Test
	if(FALSE){
		input <- lapply(testProcessed, head, n=10)
	}
	
	## Split into objects to save typing later
	review <- input$review
	business <- input$business
	user <- input$user
	checkin <- input$checkin
	
	review.a <- all$review
	business.a <- all$business
	user.a <- all$user
	checkin.a <- all$checkin
	
	## Add the information from the user
	add.user <- ldply(as.character(review$user_id), processUser, user.a=user.a, .parallel=parallel)
	
	## Process the checkin info
	add.checkin <- ldply(as.character(review$business_id), processCheckin, checkin.a=checkin.a, small=small, .parallel=parallel)
	if(small) {
		colnames(add.checkin) <- "checkin.total"
	}
	
	## Process the business info
	add.business <- ldply(as.character(review$business_id), processBusiness, business.a=business.a, .parallel=parallel)
	
	## Remove uninformative columns from the review
	colnames(review)[7:206] <- paste0("word.", colnames(review)[7:206])
	add.review <- review[, !colnames(review) %in% c("type")]
	if(small){
		del <- grep("word.", colnames(add.review))
		add.review <- add.review[, -del[51:length(del)]]
	}
	
	## Merge the results
	result <- cbind(add.review, add.user, add.business, add.checkin)
	
	## Done
	return(result)
}
```

Actually process the data and save the results
```{r runProc}
testReview <- processFun(testProcessed, all=all, small=TRUE)
trainingReview <- processFun(trainingProcessed, all=all, small=TRUE)

print(object.size(testReview), units="Mb")
print(object.size(trainingReview), units="Mb")

save(testReview, file="testReview.Rdata")
save(trainingReview, file="trainingReview.Rdata")
```


Reproducibility
```{r reproduc}
print(proc.time())
sessionInfo()
```
